Synthesis of Rules for Network Intrusion Detectors

### 2019-10-23 paper/paper.tex
pass on intro+background

\begin{table}[h]
  \caption{\label{table:rules}Example options.}  
  \centering
  \begin{tabular}{p{0.5cm}lrp{8cm}}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{\# Params} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & - & matches on the size of the packet payload\\
    2 & itype & - & matches on a specific ICMP type\\
    3 & icode & - & matches on a specific ICMP code\\
    4 & icmp\_seq & - & checks a ICMP sequence number\\
    5 & icmp\_id & - & matches on specific ICMP id-values\\
    6 & window & - & checks the size of the TCP window\\
    7 & flags & - & matches on TCP flags\\
    8 & fragbits & - & checks if the fragmentation or reserved bits are set in the IP header\\
    9 & threshold & - & controls the rule’s alert frequency\\
    10 & content & - & matches on bytes of the payload\\
    \bottomrule
  \end{tabular}
\end{table}

@@

@@

The main challenge in synthesizing rules is to discover the set of
options that makes a rule to capture the malicious traffic, but
ignores the benign traffic.

\tname{} synthesizes rules with the action header fixed to
\CodeIn{alert} and with the header \CodeIn{<proto> any any -> any
  any}, indicating that it infers the target protocol and uses the
most general flow description possible, \ie{}, the rule applies to the
traffic flowing from/to any address and port. The rule that \tname{}
synthesizes for the SYN Flood denial-of-service attack is as follows.

\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
%  \label{fig:synflood-example-synt}
\end{figure}

\Mar{Guilherme/Lucas (CRI): we need to show how \tname{} works on this
  example. What is the input? How long does it take to generate the
  rule? Is the rule generated exactly the same? Please explain...}

\begin{table}[h]
  \caption{\label{table:rules}Rule evolution}  
  \centering
  \begin{tabular}{lllll}
    \toprule
    \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
    \midrule     
    0 & (ack:0; seq:0; window:0; flags:F;)\\
    5 & (window:0; flags:F;)\\
    68 & (window:64; flags:F;)\\
    73 & (window:64; flags:S,12;)\\
    94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
    \bottomrule
  \end{tabular}
\end{table}

@@

>>>>

\begin{table}[h]
  \caption{\label{table:rules}Example options.}  
  \centering
  \begin{tabular}{p{0.35cm}ll}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & matches on the size of the packet payload\\
    2 & itype &  matches on a specific ICMP type\\
    3 & icode & matches on a specific ICMP code\\
    4 & icmp\_seq  & checks a ICMP sequence number\\
    5 & icmp\_id & matches on specific ICMP id-values\\
    6 & window & checks the size of the TCP window\\
    7 & flags & matches on TCP flags\\
    8 & fragbits & checks if the fragmentation or reserved bits are set in the IP header\\
    9 & threshold & controls the rule’s alert frequency\\
    10 & content & matches on bytes of the payload\\
    \bottomrule
  \end{tabular}
\end{table}

@@

In the following, we illustrate how \tname{} proceeds for obtaining
the rule for the SYN Flood attack. The main challenge in synthesizing
rules is to discover the set of options that enables the \nids\ to
capture the malicious traffic, but ignore the benign traffic.  The
other parts of the rule (\ie{}, the action and the header) are either
inferred or left constant. \tname{} sets the action header to
\CodeIn{alert} and sets the header to \CodeIn{<proto> any any -> any
  any}, indicating that it infers the target protocol, and it uses the
most general flow description possible, \ie{}, analyzing the traffic
flowing from/to any address and port.  We conjecture that the system
administrator would know how to modify the constant parts.

\tname{} starts the search for rule options for the SYN Flood attack
from the initial set of options appearing below.

\Fix{atualize isto para o valor correto, por favor. $\rightarrow$}
\begin{figure}[h!]
  \vspace{-3ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={flags,threshold}]{synflood.suricata.synth}
%  \caption{Suricata rule for SYN Flood Attacks.}
  %  \label{fig:synflood-example-synt}
  \vspace{-3ex}  
\end{figure}

This set is based on an analysis of the packets from the positive
traffic. Consequently, a rule that uses this set of options would
capture the positive traffic by construction. Unfortunately, such rule
would be overspecified, \ie{}, the rule would unlikely capture the
same attacks as the traffic data would not be the same and \tname{}
includes in the set all possible options that can be inferred from the
data. To address that, \tname{} minimizes the rule with the goal of
finding the minimal set of rules that only captures the positive
traffic, hence mitigating the overfitting problem. \tname{}
potentially finds several minimal rules. The list below shows the
options for the five shortest rules it produces for this attack.

\Fix{...elaborate...}

@@

\subsection{\Fix{...some other with contents...}}

\Fix{...elaborate...}

%% \begin{table}[h]
%%   \caption{\label{table:rules}Rule evolution}  
%%   \centering
%%   \begin{tabular}{lllll}
%%     \toprule
%%     \multicolumn{1}{c}{Iteration} & \multicolumn{1}{c}{Rule} \\
%%     \midrule     
%%     0 & (ack:0; seq:0; window:0; flags:F;)\\
%%     5 & (window:0; flags:F;)\\
%%     68 & (window:64; flags:F;)\\
%%     73 & (window:64; flags:S,12;)\\
%%     94 & (window:64; flags:S,12; threshold: type both, track by\_dst, count 20, seconds 1;)\\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

@@

\Fix{---------------------------------------------------
  estou aqui}
